---
title: "EDA Assignment Prep"
author: "Amber Camp"
format: html
editor: visual
---

## EDA Assignment Prep: Pokemon data

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r}
pokedata <- read_csv(here("data/Pokemon.csv"))

pokedata <- pokedata %>%
  clean_names
```

### Describe the data:

Data author: **Alberto Barradas**

Description: "This data set includes 721 Pokemon, including their number, name, first and second type, and basic stats: HP, Attack, Defense, Special Attack, Special Defense, and Speed. It has been of great use when teaching statistics to kids. With certain types you can also give a geeky introduction to machine learning.

This are the raw attributes that are used for calculating how much damage an attack will do in the games. This dataset is about the pokemon games (*NOT* pokemon cards or Pokemon Go)."

Link: <https://www.kaggle.com/datasets/abcsds/pokemon?resource=download>

How many observations (rows) are there?: 800

How many variables (columns) are there?: 13

### skimr: Quick overview

```{r}
skim_data <- skimr::skim(pokedata)
```

### Summary: Quick overview

```{r}
summary(pokedata)
```

### Categorical vs. Continuous

```{r}
class(pokedata$number) #checks class
pokedata$number <- as.factor(pokedata$number) #converts to factor, as.numeric would convert it to numeric
class(pokedata$number)
summary(pokedata$number)

summary(pokedata$type_1)
pokedata$type_1 <- as.factor(pokedata$type_1)
class(pokedata$type_1)
summary(pokedata$type_1)

summary(pokedata$generation)
```

How many categorical variables are there? Five \[maybe six\] (number, name, type_1, type_2, generation, \[legendary\]).

How many continuous variables are there? Seven (total, hp, attack, defence, sp_atk, sp_def, speed).

What are some variables of interest to you? type_1 and hp.

```{r}

```

------------------------------------------------------------------------

## EDA Assignment Prep: \[Insert Data Name Here\]

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r}

```

### Describe the data:

Data author:

Description:

How many observations (rows) are there?

How many variables (columns) are there?

### Summary: Quick overview

```{r}

```

### Categorical vs. Continuous

```{r}

```

How many categorical variables are there?

How many continuous variables are there?

What are some variables of interest to you?
