---
title: "EDA Assignment Prep"
author: "Amber Camp"
format: html
editor: visual
---

## EDA Assignment Prep: Pokemon data

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r}
pokedata <- read_csv(here("data/Pokemon.csv"))

pokedata <- pokedata %>%
  clean_names
```

### Describe the data:

Data author: **Alberto Barradas**

Description: "This data set includes 721 Pokemon, including their number, name, first and second type, and basic stats: HP, Attack, Defense, Special Attack, Special Defense, and Speed. It has been of great use when teaching statistics to kids. With certain types you can also give a geeky introduction to machine learning.

This are the raw attributes that are used for calculating how much damage an attack will do in the games. This dataset is about the pokemon games (*NOT* pokemon cards or Pokemon Go)."

Link: <https://www.kaggle.com/datasets/abcsds/pokemon?resource=download>

How many observations (rows) are there?: 800

How many variables (columns) are there?: 13

### skimr: Quick overview

```{r}
skim_data <- skimr::skim(pokedata)
#Gives context for the type of data in a dataset (in the column "skim type")
```

### Summary: Quick overview

```{r}
summary(pokedata)
```

### Categorical vs. Continuous

```{r}
class(pokedata$number) #checks class
pokedata$number <- as.factor(pokedata$number) #converts to factor, as.numeric would convert it to numeric. reason why this is being converted to factor is because the "number" is actually a label and not measuring anything numeric
class(pokedata$number)
summary(pokedata$number)

summary(pokedata$type_1)
pokedata$type_1 <- as.factor(pokedata$type_1)
class(pokedata$type_1)
summary(pokedata$type_1)

pokedata$name <- as.factor(pokedata$name)
pokedata$type_2 <- as.factor(pokedata$type_2)
pokedata$generation <- as.factor(pokedata$generation)

skim_data <- skimr::skim(pokedata)
```

How many categorical variables are there? Five \[maybe six\] (number, name, type_1, type_2, generation, \[legendary\]).

How many continuous variables are there? Seven (total, hp, attack, defence, sp_atk, sp_def, speed).

How many numeric variables are there?

What are some variables of interest to you? type_1 and hp.

------------------------------------------------------------------------

## EDA Assignment Prep: \[Insert Data Name Here\]

### Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)
library(dplyr)
```

### Data

```{r}
flowerdata <- read_csv(here("data/flowers.csv"))
```

### Describe the data:

Data author: Maryna Shut

Description: This is a dataset with information about 73 different flowers: their names, place of origin, height, longevity, industry use and number of petals.

How many observations (rows) are there? 72

How many variables (columns) are there? 10

### Summary: Quick overview

```{r}
summary(flowerdata)
```

### Categorical vs. Continuous

```{r}
skim_data_flowers <- skimr::skim(flowerdata)
```

```{r}
str(flowerdata)

categorical_vars <- flowerdata %>%
  select_if(function(col) is.factor(col) | is.character(col))

continuous_vars <- flowerdata %>%
  select_if(is.numeric)

num_categorical <- ncol(categorical_vars)
num_continuous <- ncol(continuous_vars)

cat("Number of categorical variables:", num_categorical, "\n")
cat("Number of continuous variables:", num_continuous, "\n")
```

How many categorical variables are there? 7.

How many numeric (continuous) variables are there? 3.

What are some variables of interest to you? Color and climate, Place of origin and average number of petals.

What kinds of questions could you explore with this data? Various questions and relationships can be explored with this data. Such as: "Is there a correlation between the color(s) of a flower and the climate where it grows?," or, "Does a flower's place of origin determine the number of petals it has?".
